---
title: "Lessson_7"
author: "Fernando Rodriguez"
date: "2/21/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Create lesson 7 R-Markdown<br>
2. Install following packages<br>
    - install.packages("rforcats")<br>
    - install.packages("kableExtra")<br>
    - install.packages("psych")<br>
3. Load libraries below<br>
4. Load 'Physics Course GB SV DEMO data.csv' as 'fulldata'<br>

# 0a. Load Libraries
```{r}
library(readr) # for reading in data
library(ggplot2) # for plotting
library(summarytools) # for summarizing data
library(dplyr) # for selecting and filtering
library(forcats) # for working with categorical variables
library(kableExtra) # for making tables
library(psych) # for doing descriptives and analyses
```


# 0b. Load data
```{r}
fulldata <- read_csv("/Volumes/GoogleDrive/My Drive/Intro to R Workshop - Winter 2019/Data Files/Physics Course GB SV DEMO data.csv")
```


# 1. Working with Factors/Characters/Strings

```{r}
ggplot(fulldata, aes(x = major1)) + geom_bar() + coord_flip()
```

# Using forcats() library to change the order

```{r}
str(fulldata)
```

Let's work with the major1 variable

We see that our import assigned this variable as a character. We want to change
this into a factor.

When we use summary() [BaseR], 
```{r}
summary(fulldata$major1)
```

Changing structure type to factor using as.factor() you can also use factor()
```{r}
fulldata$major1 <- as.factor(fulldata$major1)
```


```{r}
summary(fulldata$major1)
```

## Using fct_count() function

```{r}
fct_count(fulldata$major1)
```

## using fct_infreq()

```{r}
tableexample1 <- fct_count(fct_infreq(fulldata$major1))
```

## using fct_infreq() in ggplot
```{r}
ggplot(fulldata, aes(x = fct_infreq(major1))) + geom_bar() + coord_flip()
```

## use the fct_rev() function to reverse the order
```{r}
ggplot(fulldata, aes(x = fct_rev(fct_infreq(major1)))) + geom_bar() + coord_flip()
```




```{r}
fulldata$major_recode <- as.character(fulldata$major1)
fct_count(fct_infreq(fulldata$major_recode))
```

# 2/22/19 Fixing Recoding Issue
# Initialize the fulldata$major_recode variable 
```{r}
fulldata$major_recode <- as.character(fulldata$major1)
```


```{r}
# Pharmaceutical Sciences
fulldata$major_recode[fulldata$major1 == "Pharmacology and Toxicology"] <- "Pharmaceutical Sciences"

# cs
fulldata$major_recode[fulldata$major1 == "Computer Science"] <- "Computer Science"
fulldata$major_recode[fulldata$major1 == "Computer Game Science"] <- "Computer Science"
fulldata$major_recode[fulldata$major1 == "Software Engineering"] <- "Computer Science"

# arts and humanities
fulldata$major_recode[fulldata$major1 == "Dance"] <- "Arts and Humanities"
fulldata$major_recode[fulldata$major1 == "Art"] <- "Arts and Humanities"
fulldata$major_recode[fulldata$major1 == "Art History"] <- "Arts and Humanities"
fulldata$major_recode[fulldata$major1 == "History"] <- "Arts and Humanities"
fulldata$major_recode[fulldata$major1 == "Spanish"] <- "Arts and Humanities"


# social sciences
fulldata$major_recode[fulldata$major1 == "Public Health Sciences"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Psychology"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Psychology and Social Behavior"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Sociology"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Political Science"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Anthropology"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Biology/Education"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Education Sciences"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Social Ecology"] <- "Social Sciences"
fulldata$major_recode[fulldata$major1 == "Social Policy and Public Servi"] <- "Social Sciences"

# Business
fulldata$major_recode[fulldata$major1 == "Business Administration"] <- "Business"
fulldata$major_recode[fulldata$major1 == "Business Economics"] <- "Business"

# enviromental sciences
fulldata$major_recode[fulldata$major1 == "Earth System Science"] <- "Environmental Science"
fulldata$major_recode[fulldata$major1 == "Environmental Science"] <- "Environmental Science"

```



## Same plot, but with the recoded major varialbe, major_recode
```{r}
ggplot(fulldata, aes(x = fct_rev(fct_infreq(major_recode)))) + geom_bar() + coord_flip()
```


## Use kableExtra to make nice looking tables
### also using psych package to provide the summary statistics

kable() function give us a nice clean table
```{r}
kable(tableexample1)
```

## We can examine grade describes and create a table using kableExtra library

First, I need to subset the variables I want to look at. Here I am looking at the 
gradebook data
```{r}
data_ss1 <- select(fulldata, quizzes, homework, finalexam, grade_finalscore)
```

Second, I want to use the data_ss1 subset to get descriptives
We are going to use the describe() function from the psych library

```{r}
tableexample2 <- describe(data_ss1)
tableexample2
```


## Using the kable_styling() function and bootstrap_options arguments within kableExtra..
- striped
- hover
- bordered
- condensed
- full_width


```{r}
# kable(tableexample2) %>% # this is a pipe, it just tells r "then do this"
#  kable_styling(bootstrap_options = c("striped"))
```

```{r}
# kable(tableexample2) %>% # this is a pipe, it just tells r "then do this"
#  kable_styling(bootstrap_options = c("striped", "hover"))
```
```{r}
# kable(tableexample2) %>% # this is a pipe, it just tells r "then do this"
#  kable_styling(bootstrap_options = c("striped", "hover", "bordered"))
```

```{r}
# kable(tableexample2) %>% # this is a pipe, it just tells r "then do this"
#  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"))
```

```{r}
# kable(tableexample2) %>% # this is a pipe, it just tells r "then do this"
#  kable_styling(bootstrap_options = c("striped", "hover", "bordered", "condensed"), 
#                full_width = FALSE)
```


