---
title: "Lesson 3"
author: "Fernando Rodriguez"
date: "1/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Load Libraries
```{r}
library(ggplot2) # for graphing
library(summarytools) # for getting data summaries
library(readr) # for reading in data
```

# 1. Import demographic data using read.csv()
```{r}
demog <- read.csv("/Volumes/GoogleDrive/My Drive/R Workspace/R-for-Ed-Data-Science/Data/Physics Course Demo Data.csv", header = TRUE)
```

run dfSummary()
```{r}
print(dfSummary(demog), method = 'browser')
```

We need to use the na.strings argument to get correct NAs

na.string replaces values in your csv file with NAs
```{r}
demog <- read.csv("/Volumes/GoogleDrive/My Drive/R Workspace/R-for-Ed-Data-Science/Data/Physics Course Demo Data.csv", header = TRUE, na.strings = "")
```

```{r}
print(dfSummary(demog), method = 'browser')
```


# importing demographic data using read_csv()
```{r}
demog2 <- read_csv("/Volumes/GoogleDrive/My Drive/R Workspace/R-for-Ed-Data-Science/Data/Physics Course Demo Data.csv")
# View(demog2)
```

compare demog and demog2
```{r}
str(demog)
```

```{r}
str(demog2)
```

# Subsetting

subset() <br>

arguments <br>
> <br>
< <br>
>= <br>
<= <br>
== <br>
!= (is not) <br>

```{r}
demog2_ss_male <-  subset(demog2, gender == "Male")
```

I want to make a bar chart with just male and females
```{r}
demog_gender <- subset(demog2, gender == "Female" | gender == "Male")
```


# Create a bar chart using ggplot
```{r}
ggplot(demog_gender, aes(x = gender)) + geom_bar(fill = 'red', alpha = .5) + facet_grid(.~firstgenerationflag)
```


Let's exclude the MDT and NAT
I want to make a bar chart with just male and females
```{r}
demog_gender2 <- subset(demog_gender, firstgenerationflag == "N" | gender == "Y")
```

```{r}
ggplot(demog_gender2, aes(x = gender)) + geom_bar(fill = 'red', alpha = .5) + facet_grid(.~firstgenerationflag)
```


