---
title: "Lesson_6"
author: "Fernando Rodriguez, Ying Xu, Ha Nguyen"
date: "2/14/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

TO-DO Prior to Lesson<br>
1. Create lesson 6 R-Markdown<br>
2. load readr, ggplot2, summarytools, and dplyr libraries <br>
3. load merged data<br>


## 0a. Load Libraries
```{r}
library(readr)
library(ggplot2)
library(summarytools)
library(dplyr)
library(cowplot)
```

## 0b. Load our physics demographic data
importing demographic data using read_csv()
```{r}
fulldata <- read_csv("/Volumes/GoogleDrive/My Drive/R Workspace/R-for-Ed-Data-Science/Data/Physics Course GB SV DEMO data.csv")
```

# 1. Scatterplot using geom_point()
## 1a. Basic scatterplot
```{r}
theme_set(theme_grey())
ggplot(fulldata, aes(x = post_study, y = grade_finalscore)) + geom_point() 

```

## 1b. Scatterplot by group (gender)
```{r}
ggplot(fulldata, aes(x = post_study, y = grade_finalscore)) + geom_point() + facet_grid(.~ gender)
```

## 1c. Change color and point size
Single color and size
```{r}
ggplot(fulldata, aes(x = post_study, y = grade_finalscore)) + geom_point(color = "red", size = 1.5) + facet_grid(.~ gender)
```

Color based on variable value
```{r}
ggplot(fulldata, aes(x = post_study, y = grade_finalscore, color = fulltimestatus)) + geom_point(size = 1.5) + facet_grid(.~ gender)
```

## 1d. Save plot as an object
```{r}
sctplot <- ggplot(fulldata, aes(x = post_study, y = grade_finalscore, color = fulltimestatus)) + geom_point(size = 1.5) + facet_grid(.~ gender)
sctplot
```

## 1e. Add (change) label and caption
```{r}
sctplot_label <- sctplot  + labs(x = "Weekly study time", y = "Final score", caption = "Source: UCI Physics course data.", colour = "Full time status") 
sctplot_label
```

## 1f. Add title
```{r}
sctplot_title <- sctplot_label + ggtitle("Association between Reported Study Time and Final Grade") 
sctplot_title
```

## 1g. Use predefined themes
```{r}
sctplot_theme <- sctplot_title + theme_light()
sctplot_theme
## Also try theme_dark() and theme_minimal()
```

## 1h. Add a regression line
```{r}
sctplot_final <- sctplot_theme + geom_smooth(method = "lm")
sctplot_final
```

# 2. Bar chart using geom_bar()
## 2a. Basic bar chart
```{r}
ggplot(fulldata, aes(x = grade_lettergrade)) + geom_bar()
```

## 2b. Remove missing value from the chart
```{r}
fulldata_ss <- filter(fulldata, !is.na(fulldata$grade_lettergrade))
```

## 2c. Plot the bar graph with the subsetted data again
```{r}
ggplot(fulldata_ss, aes(x = grade_lettergrade)) + geom_bar()
```

## 2d. Fill the bar graphs by another variables
```{r}
barplot_color <- ggplot(fulldata_ss, aes(x = grade_lettergrade, 
                                         fill = fulltimestatus)) + geom_bar()
barplot_color
```

## 2e. Add title, labels, and captions.

```{r}
barplot_title <- barplot_color + labs(x = "Grade", y = "Count", 
                                      fill = "Full time status", caption = "Source: UCI Physic course data") + ggtitle("Letter Grade Distribution by Full/Part Time Status ")
barplot_title

```

## 2f. Change theme
```{r}
barplot_theme <- barplot_title + theme_light()
barplot_theme
```

## 2g. Remove the gridline in barcharts
```{r}
barplot_final <- barplot_theme + background_grid(major = "none")
barplot_final
```


# 3. Putting graphs into one grid and changing gridline using the package "cowplot"
https://cran.r-project.org/web/packages/cowplot/vignettes/introduction.html

## 3a. Load cowplot package
```{r}
library(cowplot)
```

## 3b. Putting two graphs in one plot
```{r}
sct_bar_plot <- plot_grid(sctplot_final, barplot_final, labels =  c("A", "B"), nrow = 2)
sct_bar_plot
```

# 4. Save plot as png file
```{r}
save_plot("cowplot.png", sct_bar_plot, base_aspect_ratio = 1.4)
```


