---
title: "Lesson_4"
author: "Fernando Rodriguez"
date: "1/31/2019"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lesson 4 
to do<br>
		1. install.packages(“dplyr”)<br>
		2. load physics demographic data<br>
		
# Table of Contents<br>
1. Revisiting Subsetting Data<br>
2. summarize() and table()<br>
3. Using dplyr() library<br>
	- filter()<br>
	- arrange()<br>
	- select()<br>
4. Recoding data<br>
5. dplyr() continued<br>
  - mutate()<br>

# 0a. Load Libraries
```{r}
library(readr)
library(ggplot2)
library(summarytools)
library(dplyr)
```


# 0b. Load our physics demographic data
# importing demographic data using read_csv()
```{r}
demog <- read_csv("/Volumes/GoogleDrive/My Drive/Intro to R Workshop - Winter 2019/Data Files/Physics Course Demo Data.csv")
```


# 1. Subsetting Data
```{r}
# get a feel for the data
view(dfSummary(demog))

demog_ss_female <- subset(demog, gender == "Female")
```


# 2. summary() and table()
```{r}
summary(demog$hsgpa) # numeric data
table(demog$gender) # character strings
```


# 3. Use dplyr() to transform our data

filter() - pick observations(e.g. students) by a specific value
```{r}
demog_ss_gender <- filter(demog, gender == "Female" | gender == "Male")
```

filter by two different criteria
```{r}
test1 <- filter(demog, gender == "Male" & ingradebookdata == "Yes")
```

```{r}
summary(demog$hsgpa)
test2 <- filter(demog, gender == "Male" & hsgpa >= 4.0)
```

Exercise

1. create a filter function
2. that filters by students who are low-income AND first-generation students
3. Assign this filter to a new dataframe

```{r}
test3 <- filter(demog, lowincomeflag == "Y" & firstgenerationflag == "Y")
```


# agrrange()
reorder rows
```{r}
test4_arrange <- arrange(demog, major1)
```

using is.na
```{r}
x <- NA
is.na(x)
!is.na(x)

```


```{r}
demog_ss <- filter(demog, !is.na(major1))
```

# Recoding data
let's use demog_ss from now on

```{r}
demog_ss$gender_num <- NA
summary(demog_ss$gender_num)

demog_ss$gender_num[demog_ss$gender == "Male"] <- 0
demog_ss$gender_num[demog_ss$gender == "Female"] <- 1
demog_ss$gender_num[demog_ss$gender == "Not stated"] <- NA

table(demog_ss$gender_num)

table(demog_ss$gender, demog_ss$gender_num)
```

# hold down alt when selecting multiple rows to mass-edit code
```{r}
# 
# demog_ss$gender_num[demog_ss$gender == "Male"] <- 0
# demog_ss$gender_num[demog_ss$gender == "Female"] <- 1
# demog_ss$gender_num[demog_ss$gender == "Not stated"] <- NA
# 
# demog_ss$gender_num2[demog_ss$gender2 == "Male"] <- 0
# demog_ss$gender_num2[demog_ss$gender2 == "Female"] <- 1
# demog_ss$gender_num2[demog_ss$gender2 == "Not stated"] <- NA
# 
# demog_ss$gender_num3[demog_ss$gender3 == "Male"] <- 0
# demog_ss$gender_num3[demog_ss$gender3 == "Female"] <- 1
# demog_ss$gender_num3[demog_ss$gender3 == "Not stated"] <- NA
# 
# demog_ss$gender_num4[demog_ss$gender4 == "Male"] <- 0
# demog_ss$gender_num4[demog_ss$gender4 == "Female"] <- 1
# demog_ss$gender_num4[demog_ss$gender4 == "Not stated"] <- NA


```


