---
title: "Lessson_7"
author: "Fernando Rodriguez"
date: "2/21/2019"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Create lesson 8 R-Markdown
2. Install following packages
  - install.packages("rforcats")ß
  - install.packages("kableExtra")
  - install.packages("psych")
3. Load libraries below
4. Load 'Physics Course GB SV DEMO data.csv' as 'fulldata'

# 0a. Load Libraries
```{r}
library(readr) # for reading in data
library(ggplot2) # for plotting
# library(summarytools) # for summarizing data
library(dplyr) # for selecting and filtering
# library(forcats) # for working with categorical variables
# library(kableExtra) # for making tables
# library(psych) # for doing descriptives and analyses
```


# 0b. Load data
```{r}
fulldata <- read_csv("/Volumes/GoogleDrive/My Drive/Intro to R Workshop - Winter 2019/Data Files/Physics Course GB SV DEMO CLICK Data.csv")
```

# Vectors
a basic data structure containing elements of the same type.
Vectors are one dimensional arrays.

# using Vectors

We have focused on the tidyverse way to select and subset our rows and columns.

However the baseR way of doing this uses, what we call vectors.

Two types of vectors
Atomic vectors: 
logical, integer, double (double precision floating point numbers, character, complex, and raw
atomic vectors are homogeonous (they are the same type)

Lists (sometimes called recursive vectors) because lists can contain other lists
Lists are heterogenous (can be different atomic vectors)

We cam make a list using the baseR list() function
```{r}
list1 <- list(1, 4, 6, 7, 8, 9)
```



NULL is used to represent absence of a vector and is often treated as a lenghth 0
vector


Vectors have 2 key properties

1. types (what type of vector it is)
We can see this by using typeof()



# the default storage for integers are the double type
```{r}
typeof(353)

typeof(1)

typeof(1.3)
```


```{r}
typeof(fulldata$tot_click)
```


```{r}
typeof(fulldata$gender)
```

```{r}
typeof(list1)
```

2. length
how long a vector is
```{r}
length(fulldata$tot_click)
```

We see that the length of each numeric value is the independent values in a
vector or list
```{r}
length(1.3)
```

```{r}
length(list1)
```

# remember that lists can be combinations of numeric and string character values
```{r}
list2 <- list(2, "hello", 2.2)

length(list2)
```


```{r}
length(fulldata$tot_click)
```


## Character Vectors
For characters, the length is 1 for every enclosed unit "" 
```{r}
char1 <- ("hello how are you")
typeof(char1)

length(char1)
```


This is the same for vectors of characters
```{r}
typeof(fulldata$gender)

length(fulldata$gender)
```


# Naming Vectors

You can actually name vectors
```{r}
list3 <- c(x = 1, y = 2)

list3
```

subsetting vectors

We used the filter() function in dplyr (deeplyer) 

[ is the subsetting function,

and is use like this 
x[a]

There are 4 types of things you can subset vectors with

# First, lets create a vector of strings
c() is a funciton that returns a vector (a one dimensional array)
```{r}
x <- c("one", "two", "apple", "four")

typeof(x)
```

We can filter out specific characters in our vector x
each character in our vector x is indexed, starting at 1

```{r}
x[c(1, 3)]
```

you can repeat positions to make longer outputs as well
```{r}
x[c(1, 3, 4, 1, 1, 2, 2, 2, 2)]
```


negative values drop the lements at the specified position
```{r}
x[c(-1, -3)]
```

You cannot mix positive and negative subsetting values
```{r}
# x[c(-1, 3)]
```

```{r}

```



Piping
%>%

shortcut method for performing data filtering, selecting, etc.

%>% basicially states 'then do this'
try to follow the logic of how you already work with data
```{r}
fulldata %>%
  filter(gender == "Male") %>%
  ggplot(aes(x = fulltimestatus, y = grade_finalscore)) + geom_boxplot()
```

