---
title: "R Workshop for Education Data Science Lesson 1"
author: "Fernando Rodriguez"
date: "1/10/2019"
output: rmarkdown::github_document
---
<br>

## Table of Contents
1. Loading Libraries
2. Creating and running code-chunks in R-Markdown
3. The basics
    - Simple calculations
    - Assignment
4. Exploring Data
    - head() 
5. Graphic Data
    - plot()
6. A Note on Arguments
7. Intro to ggplot2 library and functions
    - aes()
    - geom()
    - theme()
    - labs()
    - scale_color_gradient()
    - facet_grid()
8. Learning More About Libraries and Functions
  


## 1. Loading Libraries
First, let's load the libraries you will use for this lesson. This is the first
thing you should do when writing an R-Markdown document. That way, you ensure
that you load all of the necessary libraries prior to running code
```{r}
# install.libraries("gglot2") <- installing libraries require double qoutations ""
library(ggplot2) # <- loading libraries doesn't require double qoutations
```
<br><br>

Before going into the ins-and-outs of coding in R, let's get familiar with
how we can use R-markdown to create chunks of code.<br>

## 2. Creating code chunks in R-Markdown
alt + command + i (mac)<br>
control + alt + i (windows)<br>

Each chunk of code can be run individually.
```{r}
a <- 3 + 3
```

### Running chunnks
You can run the chunk of code by clicking on the green arrow to the right
of the code. 

Or use theses shortcuts<br)
command + enter (mac)<br>
control + enter (windows)
<br><br>
## 3. The very basics
### Simple Caluations
```{r}
2 + 2
```

```{r}
10/2
```

### Assignment
We assign values to variables by using `<-`<br>
Assigning a, b, and c
```{r}
a <- 2 + 2
b <- 10 + 2
c <- 4 + 12
```

## Creating d, which is a + b
```{r}
d <- a + b
d
```
<br><br>

## 4. Exploring Data<br>
### `mtcars` data
This data is pre-installed in R, and provides a quick way to fiddle around
with coding.

We can see the data by simply typing `mtcars`
```{r}
mtcars
```


If you want to see less rows you can use the function `head`
```{r}
head(mtcars)
```


### `pressure` data
We can see that there are only two variables in the pressure data
```{r}
head(pressure)
```
<br><br>

## 5. Graphing Data
We can use the `plot` function to create a scatter plot for the `pressure` data 

```{r}
plot(pressure)
```
<br><br>

## 6. A Note on Arguments 

Notice that functions in R always have () beside them<br>
head(mtcars)<br>
plot(pressure)<br>
<br>
In R, we put our arguments (which are extra things we want the argument to do)
inside these parentheses.

<br><br>
## 7. Intro to ggplot library and functions
Remember to make sure ggplot is loaded into R.<br>
You can do this by running the first chunk of this document, where it has the
code `library(ggplot2)`<br>
<br>
We want to plot how miles per galon `mpg` is related to horsepower `hp`.
And then we want to split this up by cylnders `cyl`.<br><br>

for ggplot, our first argument will be the dataset `mtcars`
```{r}
ggplot(mtcars)
```
<br><br>

#### aes() function for stating your x and y axis
Within the ggplot() function, we'll set up our parameters by using the `aes()` function
aes stands for asthetic. For this function, we want to define the x and y axis.

The x-axis will be `hp` and the y-axis will be `mpg`
```{r}
ggplot(mtcars, aes(x = hp, y = mpg))
```
<br><br>

### The power of `+`
### Adding features to your ggplot graph using `+`

We can add new features by using other functions that are part of the ggplot 
library.

We do this by using the `+` sign

<br><br>
####  geom() function for stating the kind of graph you want

geom stands for geometric unit.<br>

Now let's use the `geom` family of functions to state what kind of graph we want.<br>

We want a scatterplot, so we are going to use the function `geom_point()`<br>

No arguments are required for `geom_point()`
```{r}
ggplot(mtcars, aes(x = hp, y = mpg)) + 
 geom_point() 
```
<br><br>

####  theme() function for modifying components of your graph

let's set the size of the text by 20 using `text = element_text(size = 20)`
```{r}
ggplot(mtcars, aes(x = hp, y = mpg)) + 
  geom_point() + 
  theme(text = element_text(size = 20))
```
<br><br>

#### labs() function for labeling your graph

```{r}
ggplot(mtcars, aes(x = hp, y = mpg)) + 
 geom_point() + 
  theme(text = element_text(size = 20)) +
  labs(x = "Horsepower", y = "MPG")
```
<br><br>

#### scale_color_gradient() function for using a color gradient on mpg

We want the low mpg to be blue and the high mpg to be red.<br>

IMPORTANT! In order to make this function work, you have to state which variable
you want it to color. Let's color `mpg`. <br>

In order to state this, we have to go back to the `aes()` function and write an additional argument. Remember, arguments are separated by `,`<br>

So your aes code should look like this now `aes(x = hp, y = mpg, color = mpg)`
```{r}
ggplot(mtcars, aes(x = hp, y = mpg, color = mpg)) + 
 geom_point() + 
  theme(text = element_text(size = 20)) +
  labs(x = "Horsepower", y = "MPG") +
  scale_color_gradient(low = "red", high = "blue")
```
<br><br>

#### facet_grid() function splitting up the graph by a group

We want to split our graph up by the variable `cyl`<br>

use `.~cyl` to get a horizontal verison 
```{r}
ggplot(mtcars, aes(x = hp, y = mpg, color = mpg)) + 
 geom_point() + 
  theme(text = element_text(size = 20)) +
  labs(x = "Horsepower", y = "MPG") +
  scale_color_gradient(low = "red", high = "blue") +
  facet_grid(.~cyl)
```
<br><br>

use `cyl~.` to get a vertical verison.

```{r}
ggplot(mtcars, aes(x = hp, y = mpg, color = mpg)) + 
 geom_point() + 
  theme(text = element_text(size = 20)) +
  labs(x = "Horsepower", y = "MPG") +
  scale_color_gradient(low = "red", high = "blue") +
  facet_grid(cyl~.)
```
<br><br>

### 8. Learning More About Libraries and Functions

If you want to see more information about what you can do with a library like
ggplot2, you can put `?` in front of the name of the library.

```{r}
?ggplot2
```
<br><br>
If you want to know more about how to use a specific function
put a ? in front of the function name
```{r}
?ggplot()
```
<br><br>
You can even do this with sub-functions, like `element_text()`
```{r}
?element_text()
```


