---
title: "Lesson_5"
author: "Fernando Rodriguez"
date: "2/7/2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create lesson 5 R-Markdown
2.load readr, ggplot2, summarytools, and dplyr libraries 
3. load physics demographic data
4. Save survey and gradebook data to your current data folder



# 0a. Load Libraries
```{r}
library(readr)
library(ggplot2)
library(summarytools)
library(dplyr)
```


# 0b. Load our physics demographic data
# importing demographic data using read_csv()
```{r}
demog <- read_csv("/Volumes/GoogleDrive/My Drive/Intro to R Workshop - Winter 2019/Data Files/Physics Course Demo Data.csv")
```

# 0c. Load the survey and gradebook data
```{r}
survey <- read_csv("/Volumes/GoogleDrive/My Drive/Intro to R Workshop - Winter 2019/Data Files/Physics Course SV and GB Data.csv")
```


# Selecting columns with select()

```{r}
ss <- select(survey, roster_randomid, post_ts)
```


1. Find 4 choice variables from the demog dataframe that you want to work with.

2. Use the select() function to subset the data 
    put the 4 variable in this dataframe
    
3. Include roster_randomid variable


```{r}
vars <- select(demog, roster_randomid, status, gender, agegroup, hsgpa)
```


```{r}
gender_ss <- filter(demog, gender == "Male" | gender == 
                      "Female")
```


# Merge Data
merge()

```{r}
merged <- merge(demog, survey, by = "roster_randomid", all = TRUE)
```



```{r}
merged2 <- merge(gender_ss, survey, by = "roster_randomid", all = FALSE)
```


# Data Exploration

1. Merge your demog subsetted dataframe with the survey dataframe

2. Start exploring relationships with the demog variable and the survey and/or gradebook variables

table(), summarize(), dfSummary(), ggplot(), plot() to explore potential group differences or correlations

`
#### Subsetting by male and female only
```{r}
demog_ss_gender <- filter(demog, gender == "Male" | gender == "Female")
```

#### Merging the demog_ss_gender subsetted data with the new survey (survey and gradebook) data
```{r}
merged_ss_gender <- merge(demog_ss_gender, survey, by = "roster_randomid", all = FALSE)
```

#### Plotting differences between males/females on 
post_study and grade_finalscore
```{r}
ggplot(merged_ss_gender, aes(x = post_study, y = grade_finalscore)) + geom_point() + facet_grid(.~ gender)
```


#### Adding model line with standard errors
```{r}
ggplot(merged_ss_gender, aes(x = post_study, y = grade_finalscore)) + geom_point() + facet_grid(.~ gender) + geom_smooth()
```


#### Adding regression line
```{r}
ggplot(merged_ss_gender, aes(x = post_study, y = grade_finalscore)) + geom_point() + facet_grid(.~ gender) + geom_smooth(method = "lm")
```


